---
title: "MATH208 Final Project"
author: "Barry Chen"
date: "2022-12-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r reading data}
cpu_gpu_data<-read_csv("/Users/cccxxx/Downloads/chip_dataset.csv",show_col_types = FALSE)
names(cpu_gpu_data)
```

## TASK 1

### a.

<br>

#### Analyze NA value

```{r Analyze NA value}
apply(is.na(cpu_gpu_data),2,sum)

cpu_gpu_data %>% summarise_all(list(~sum(is.na(.)))) %>%
pivot_longer(cols=everything(),names_to = "Variable")

```

There are 625 missing values for TPD measures, 715 missing values for Die size measures, 711 missing values for Transistors measures, and 0 missing value for Freq measures.

<br>

#### Overall numerical summary of mean and median for each measures

```{r overall numerical summary for mean and median,message=FALSE}
# Find the group size for TYPE
cpu_gpu_data <- cpu_gpu_data %>% mutate(TYPE = factor(Type))
cpu_gpu_data %>% group_by(TYPE) %>% count()
# Include the NA value
cpu_gpu_data %>% group_by(TYPE) %>% select(`Process Size (nm)`,`TDP (W)`,`Die Size (mm^2)`,`Transistors (million)`,`Freq (MHz)`) %>% summarise_all(list(Avg=mean,Med=median))

# Exclude the NA value and pivot longer
cpu_gpu_data %>% group_by(TYPE) %>% select(`Process Size (nm)`,`TDP (W)`,`Die Size (mm^2)`,`Transistors (million)`,`Freq (MHz)`) %>% summarise_all(list(Avg=mean,Med=median),na.rm = TRUE) %>% pivot_longer(cols = c("Process Size (nm)_Avg":"Freq (MHz)_Med"),names_to = "Measure")

# Exclude the NA value and pivot wider
cpu_gpu_data %>% group_by(TYPE) %>% select(`Process Size (nm)`,`TDP (W)`,`Die Size (mm^2)`,`Transistors (million)`,`Freq (MHz)`) %>% summarise_all(list(Avg=mean,Med=median),na.rm = TRUE) %>% pivot_longer(cols = c("Process Size (nm)_Avg":"Freq (MHz)_Med"),names_to = "Measure") %>% pivot_wider(id_cols = "Measure",names_from = "TYPE") %>% arrange(desc(Measure))

```

From the last table which generated by pivot wider, it's easy to summarize:

-   There are 2192 numbers of observations for CPU and 2662 numbers of observations for GPU.

-   GPU has a larger value on Transistors.

-   GPU has a larger average value on TDP.

-   GPU has a larger value on Process Size.

-   CPU has a significant larger value on Freq. GPU has a larger value on Die Size.

<br>

#### Graphically distribution of Process Size (nm)

```{r graphically distribution of Process Size (nm), message=FALSE}
# summary table
process_time <- cpu_gpu_data %>% group_by(TYPE,`Process Size (nm)`) %>% summarise(n = n())
process_time
# histogram
ggplot(cpu_gpu_data, aes(x=`Process Size (nm)`,group=TYPE,fill=TYPE)) +
geom_histogram(bins=35,col="black",na.rm = FALSE) + facet_wrap(~TYPE)

ggplot(cpu_gpu_data, aes(x=`Process Size (nm)`,group=TYPE)) +
geom_histogram(aes(y=..density..,fill=TYPE), bins=35,col="black",alpha=0.4,na.rm = FALSE) + geom_density(size=0.65,adjust=0.4) +
facet_wrap(~TYPE)
# box plot
ggplot(cpu_gpu_data,aes(x=TYPE,y=`Process Size (nm)`)) +
stat_boxplot(geom="errorbar",width=0.25) + geom_boxplot() +
ylab("Process Size (nm)")
```

From both histogram and box plot, the GPU seems have a larger mean value of process size. They both have a positive skew. But CPU seems have a wider spread on process size. Furthermore, GPU has an outlier showing in the the box plot, but CPU does not.

<br>

#### Graphically distribution of TDP (W)

```{r graphically distribution of TDP (W),message=FALSE}
TDP <- cpu_gpu_data %>% group_by(TYPE,`TDP (W)`) %>% summarise(n = n())
TDP

ggplot(cpu_gpu_data, aes(x=`TDP (W)`,group=TYPE,fill=TYPE)) +
geom_histogram(bins=35,col="black",na.rm = FALSE) +
facet_wrap(~TYPE)

ggplot(cpu_gpu_data,aes(x=TYPE,y=`TDP (W)`)) +
stat_boxplot(geom="errorbar",width=0.25) + geom_boxplot() +
ylab("TDP (W)")
```

In term of TDP, CPU and GPU have a really similar central location, i.e, their mean and median are very close. They are both positive skew and have a narrow spread on TDP. GPU seem has a wider spread for central data than CPU. GPU has a lot outliers and spread widely toward large TDP compared to CPU.

<br>

#### Graphically distribution of Die Size (mm\^2)

```{r graphically distribution of Die Size (mm^2),message=FALSE}
die_size <- cpu_gpu_data %>% group_by(TYPE,`TDP (W)`) %>% summarise(n = n())
die_size

ggplot(cpu_gpu_data, aes(x=`Die Size (mm^2)`,group=TYPE,fill=TYPE)) +
geom_histogram(bins=35,col="black",na.rm = FALSE) +
facet_wrap(~TYPE)

ggplot(cpu_gpu_data,aes(x=TYPE,y=`Die Size (mm^2)`)) +
stat_boxplot(geom="errorbar",width=0.25) + geom_boxplot() +
ylab("Die Size (mm^2)")
```

In terms of Die Size, there is no obvious differences between the median and mean. They are both positive skew. The die size of GPU spread relative wider than CPU, and it has more outliers than CPU. <br>

#### Graphically distribution of Transistors (million)

```{r graphically distribution of Transistors (million),message=FALSE}
transistors <- cpu_gpu_data %>% group_by(TYPE,`TDP (W)`) %>% summarise(n = n())
transistors

ggplot(cpu_gpu_data, aes(x=`Transistors (million)`,group=TYPE,fill=TYPE)) +
geom_histogram(bins=35,col="black",na.rm = FALSE) +
facet_wrap(~TYPE)

ggplot(cpu_gpu_data,aes(x=TYPE,y=`Transistors (million)`)) +
stat_boxplot(geom="errorbar",width=0.25) + geom_boxplot() +
ylab("Transistors (million)")
```

For Transistors, the spread of CPU and GPU are both very narrow. They have similar central location, and both positive skewed. But GPU seems has more and extreme outliers than CPU.

<br>

#### Graphically distribution of Freq (MHz)

```{r graphically distribution of Freq (MHz),message=FALSE}
freq <- cpu_gpu_data %>% group_by(TYPE,`Freq (MHz)`) %>% summarise(n = n())
freq

ggplot(cpu_gpu_data, aes(x=`Freq (MHz)`,group=TYPE,fill=TYPE)) +
geom_histogram(bins=35,col="black",na.rm = FALSE) +
facet_wrap(~TYPE)

ggplot(cpu_gpu_data,aes(x=TYPE,y=`Freq (MHz)`)) +
stat_boxplot(geom="errorbar",width=0.25) + geom_boxplot() +
ylab("Freq (MHz)")
```

For Freq, there is a clear evidence that CPU has a larger mean and median than GPU. CPU looks very like a normally distribution and spread widely, but GPU is positive skewed and spread narrowly. Also, GPU has more outlier than CPU.

<br>

#### b.

<br>

#### Association between the number of processors released by the vendors and the foundries

```{r association between the number of processors released by the vendors and the foundries,message=FALSE}

cpu_gpu_data %>% group_by(Vendor,Foundry,TYPE) %>% summarise(n=n())

cpu_gpu_data %>% group_by(Vendor,Foundry,TYPE) %>% summarise(n=n()) %>% pivot_wider(id_cols = c(Vendor,Foundry), names_from = TYPE, values_from = n)

cpu_gpu_data <- cpu_gpu_data %>% mutate(foundry_lump = fct_lump(Foundry,7))
ggplot(cpu_gpu_data,aes(x=Vendor,fill= foundry_lump)) +
geom_bar(position="dodge")  + scale_fill_viridis_d() + facet_grid(Vendor~TYPE)

```

From the wider summary table in which one can see the numerical summary of the number of chip processor released across vendor and foundry grouped by CPU and GPU and the bar plot that has foundry lump collapsed into seven levels, one can find that for Vendor of AMD, it release semiconductors nearly exclusively from unknown foundries for CPU, but release most semiconductors from TSMC. Thus, there is an association between the number of processors released but it depends on the type. For ATL, its GPU semiconductors are almost all released from TSMC, and doesn't depend on type. For Intel, no matter which type, it released exclusively from Intel itself. For NVIDA, It released almost exclusively from TSMC, and doesn't depend on type. Overall, I would like to conclude that there is a strong association between the number of processors released by the vendors and the foundries, and doesn't depend on much whether they are CPU or GPU.

<br>

#### c.

<br>

#### Association between die size and TDP

```{r asscociation between die size and TDP}
corrolation = vector("numeric",2)
names(corrolation) = c("CPU","GPU")

data_split <- with(cpu_gpu_data,split(cpu_gpu_data,TYPE))
cor1 = with(data_split[[1]],cor(`Die Size (mm^2)`,`TDP (W)`,use="complete.obs"))
corrolation[1] = cor1

data_split <- with(cpu_gpu_data,split(cpu_gpu_data,TYPE))
cor2 = with(data_split[[2]],cor(`Die Size (mm^2)`,`TDP (W)`,use="complete.obs"))
corrolation[2] = cor2
print(corrolation)

ggplot(cpu_gpu_data,aes(x=`Die Size (mm^2)`,y=`TDP (W)`,col=TYPE)) + 
geom_point() + facet_wrap(~TYPE) + geom_smooth(method="lm",col="black")

```

I computed the correlations between die size and TPD using `cor()` function and split the result according to the type. You can see the result that there is a stronger correlation for GPU between die size and TDP, but there is also a positive correlation for CPU between die size and TDP. But from the graph, there is more larger value of TDP as die size going larger for GPU. Therefore, It seems that this correlation is dependent on the type.

\newpage

## TASK 2
